Stack Safety as a Security Property

### 2020-07-07 main.tex
Pass though 2-4

But what {\em is} this ``stack safety'' of which we speak? Precise formal
definitions are lacking in the literature; instead, the notion of safety
is usually defined negatively, by enumerating the bad things that a safe system
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
We are only aware of one, quite recent, attempt to frame a positive definition.
\citet{Skorstengaard+19} define stack safety as
the conjunction of two properties, {\em local state encapsulation} and {\em
  well-bracketed control flow}.
They formalize these properties by defining an abstract capability-based machine
that ``satisfies [stack safety] by construction.''
%
While this is an important step in the right direction, we believe
there is still room to craft a more abstract, generic characterization
of stack safety that is independent of any particular enforcement
mechanism.  In other words, we seek formal criteria that can be applied
against \emph{any} \apt{stack-based?} system to determine whether or not
the system is stack-safe.  
\apt{another stab.}
\leo{I took a stab. I think it's still to dense, but captures the
  feeling of today's discussion I think}
\apt{still needs tuning}\bcp{Yes, I think we need to go into a
  little more detail on what their definition looks like in order to
  justify why an alternative definition might be valuable.  What
  questions does their presentation not answer? What uses does it not
  support? E.g., can it be enforced efficiently? (I guess so, by
  capabilities, though I think their capabilities are more powerful
  than what Cheri actually provides, while our policies are exactly
  what Nick needs.)  Why would it be good to look at stack safety
  through the lens of security?  (Because it is fundamentally a
  security mechanism?)}  \rb{BCP idea: contrast intrinsic
  vs. extrinsic property definition.  Remember to talk about spatial
  vs temporal safety}

The main contribution of this paper is a novel formal presentation of
local state encapsulation and well-bracketed control flow using the
conceptual and technical framework of language-based security.
%
The underlying notion behind local state encapsulation is the intention to
protect callers from callees: a callee should be not be able to access or
overwrite private data in the caller's stack frame. We formalize this
intuition as a security property with both confidentiality and integrity
components.
%
Intuitively, {\em stack confidentiality} promises that a callee's
observable behavior is not affected by memory outside its stack frame
(e.g. a caller's private data or uninitialized memory).  {\em Stack
  integrity} guarantees that, if a function call ever returns, the caller's
stack data at the point of return will be identical to what it was before
the call.  \bcp{We need to point out somewhere that we are only dealing with
stack safety, though we discuss (somewhere) how our property can be combined
with the heap safety property of [Azevedo et al]}\leo{Should this be
  here or in the assumptions as presented in the running example?}

@@

To provide evidence that our definitions capture the inituitive notion
of stack safety, we show that they rule out a range of exploitable bad
behaviors.\apt{``from the literature''??}  To assess whether they are
implementable, we compare them against 
%proposed capability-based
%enforcement mechanism of \citet{Skorstengaard+19} \apt{we really
%  should} and
the existing tag-based enforcement mechanism of
\citet{DBLP:conf/sp/RoesslerD18}.
%\leo{Is this where we would rephrase this because we would only talk about it in related work?}

In summary, we offer the following contributions:
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
  Abstract formal properties capturing the
  intuitive concepts of strict local state encapsulation and well-bracketed
  control flow~(\cref{sec:lse-and-wbcf}) in a generic setting. These properties
  rule out common examples of stack memory abuse.
\item
  A case study of the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\cref{sec:enforcement}).
  This policy enforces our strict stack safety property, and also a (provably) stronger property that is easier to test.
  \apt{If we have no testing, do we even want to talk about the testing
    property?} \bcp{Not here; just in future work.  I guess we need to see
    whether we wind up with any results at all about testing.}
\item
  A refinement of the case study that considers the more efficient Lazy Tagging and Clearing policy
  of \citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:lazy}).  This policy does not enforce our strict
  local state encapsulation property,
  but a modified version of it using unique per-activation tags enforces 
  a (provably) weaker lazy variant of the property. 
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}

@@

The example program appears in \Cref{fig:running-program}. It consists of a
{\tt main} function, written in some C-like high-level language, that
performs a call to some function {\tt f}, which in turns makes a nested call
to some function {\tt g} before returning, performing some simple arithmetic
in the process.

@@

\newcounter{pcctr}
\setcounter{pcctr}{0}
\newcommand*{\row}[4]{
  \thepcctr & \stepcounter{pcctr} #1 & #2 & #3 & \text{#4}\\
}
\newcommand*{\tracerow}[5]{
  #1 & #2 & #3 & #4 & \text{#5}\\
}
\newcommand*{\summary}[6]{
\[
\begin{array}{cccccc}
  \PCname = #1 & r_0 = #2 & \rra = #3 & \rsp = #4 & r_4 = #5 & r_5 = #6
\end{array}
\]
}
\newcommand{\negate}{\textrm{-}}

\newcommand{\instrc}{magenta}
\newcommand{\outc}{white}
\newcommand{\fprivc}{cyan}
\newcommand{\fcallc}{green}
\newcommand{\gprivc}{yellow}
\newcommand{\gcallc}{orange}
\newcommand{\hprivc}{red}
\newcommand{\unusedc}{gray}

@@

  \begin{array}{c|c|c|c|c}
    PC & \mathit{Instruction} & \mathit{Description} & \mathit{Effect} & \\

@@

        {Make the call}

@@

  \apt{How's that description for last line?}
  \apt{Where did these RISC-V insruction formats come from? I think some of them are just wrong, e.g.
    $\jal$ takes a PC-relative argument, not an absolute one, and the proper way to code a return to $\rra$ is $\jalr  ~ \rra ~ r_0$.
  But also the instruction operand ordering is awkward. Why not use standard AT\&T syntax?}

@@

  \begin{array}{c|c|c|c|c}

@@

    \row{\jal ~ \rra ~ 40}{??}{\PCname \leftarrow 40; ~ \rra \leftarrow \thepcctr}
        {Make the call to {\tt h}}

@@

\begin{figure}
\[
\begin{array}{c|c|c|c|c}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp - 2] \leftarrow r_4}{[107] \leftarrow 18}
      {Store $r_4$ as the return value}
  \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\rra \leftarrow 27}
      {Load return address}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 1}{\rsp \leftarrow \rsp - 1}{\rsp \leftarrow 108}
      {Restore the stack pointer of {\tt g}}
  \row{\jalr ~ \rra ~ \rra ~ 0}{\leo{??}}{\PCname \leftarrow 27}
      {Return to {\tt g}}
  \end{array}
\]

@@

\setcounter{pcctr}{27}
% H - Body
\vspace*{0.2em}
\[
\begin{array}{c|c|c|c|c}
  \row{\lw ~ r_4 ~ \rsp ~ 0}{r_4 \leftarrow [\rsp]}{r_4 \leftarrow 18}
      {Load return value to $r_4$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[105] \leftarrow 18}
      {Set {\tt w} to 18}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 106}
      {Deallocate {\tt h}'s call frame}
  \row{\lw ~ r_5 ~ \rsp ~ \negate 4}{r_5 \leftarrow [\rsp - 4]}{r_5 \leftarrow 0}
      {Load argument {\tt z} $r_5$}
  \row{\sw ~ \rout ~ r_5 ~ 0}{[\rout] \leftarrow r_5}{[99] \leftarrow 0}
      {Print {\tt z}}
  \row{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow r_4 + r_5}{r_4 \leftarrow 18}
      {Add $r_4$ and $r_5$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[103] \leftarrow 18}
      {Store $r_4$ as ther return of {\tt g}}
  \row{\lw ~ \rra ~ \rsp ~ \negate 2}{\rra \leftarrow [\rsp - 2]}{\rra \leftarrow 8}
      {Load return address}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 104}
      {Deallocate {\tt g}'s local state}
  \row{\jalr ~ \rra ~ \rra ~ 0}{??}{PC \leftarrow 8}{Return to {\tt main}}
\end{array}
\]
% Return to main
% Memory Layout
\begin{center}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{43.5em}{2em}{\SP}
\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{white}[{\makebox[0pt]{Output}}]%
\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
\memory{2}{\fcallc}[{\makebox[0pt]{{\tt g}'s call frame }}]%
\memory{6}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-51em}{0.75em}{0}
\MemoryLabel{-47em}{0.75em}{42}
\MemoryLabel{-43em}{0.75em}{0}
\MemoryLabel{-39em}{0.75em}{0}
\MemoryLabel{-35em}{0.75em}{18}
\\
\end{center}
% Register File
\summary{8}{0}{\thepcctr}{104}{18}{0}
%\vspace*{0.2em}
\setcounter{pcctr}{8}
\[
\begin{array}{c|c|c|c|c}
  \row{\lw ~ \rsp ~ r_4 ~ \negate 1}{r_4 \leftarrow [\rsp - 1]}{r_4 \leftarrow 18}
      {Load return value to $r_4$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[101] \leftarrow 18}
      {Set {\tt y} to the result}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 4}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 102}
      {Deallocate {\tt g}'s call frame}
  \row{\lw ~ \rsp ~ r_5 ~ \negate 2}{r_5 \leftarrow [\rsp - 2]}{r_5 \leftarrow 42}
      {Load {\tt x} to $r_5$}
  \row{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow r_4 + r_5}{r_4 \leftarrow 60}
      {Add {\tt x} and {\tt y}}
  \row{\sw ~ \rout ~ r_4 ~ 0}{[\rout] \leftarrow r_4}{[99] \leftarrow 60}
      {Print {\tt x + y}}
\end{array}
\]
\vspace*{0.2em}
% Return to main
% Memory Layout
\begin{center}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{31.5em}{2em}{\SP}
\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{white}[{\makebox[0pt]{Output}}]%
\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
\memory{8}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-51em}{0.75em}{60}
\MemoryLabel{-47em}{0.75em}{60}
\MemoryLabel{-43em}{0.75em}{42}
\\
\end{center}
% Starting Register File
\summary{\thepcctr}{0}{37}{102}{60}{42}
\caption{Detailed Execution Trace
  \rb{A suggestion for the description of $\jal$ and $\jalr$:
    call(function label) [if labels are added next to the entrypoints]
    or call(address of entrypoint) and return()?}
}
\label{fig:running-trace-b}
\end{figure}

To show a trace of the execution of this program, we need
a low-level view of both the code and the machine it runs on. While
our concrete implementation (Section~\leo{cite}) operates on a full
conventional RISC-V architecture, we make a few simplifying
assumptions here to ease the exposition.

\bcp{We should also emphasize, again, that there is no heap (if there were,
we might put it above the stack or between the instructions and the stack)}
First of all, we ignore the heap. We are mainly concerned
with stack interactions so that will be the main focus. \leo{We're
  going to come back to that at the end. Will we?} We also
abstract away low-level details of memory layout and present a
machine that is word-addressed, with instructions starting from
location 0, a dedicated output at location 99, and lay the stack from
location 100 upwards. \leo{Explain more? How to connect to realistic
  one? Observations? DMA? Interrupt handlers?}
%
A sketch of the memory layout at that point can be found in the top of
\cref{fig:running-trace-b}: instructions for {\tt main} begin at
0, the dedicated output is empty\apt{what does that mean? why not just uninitialized?}, while the (uninitialized) stack
begins at 100. Directly below the layout, we summarize the parts of
the register file that are relevant to the example.
%
The rest of \cref{fig:running-trace-b} details the execution of the
program, showing the machine state at call and return points, as well
as describing the full effect of each instruction in detail.
\leo{Expand upon this?}

@@

\sna{Variations here collides with state variations. I propose substitutions.}

\leo{TODO: Make attacker model more explicit} The execution trace
shown in \cref{fig:running-trace-b} is an example of a
``stack-safe'' execution: each process interacts only with the
registers, its local stack space and its call frame. But what could go
wrong if an attacker could hijack an arbitrary call and execute
arbitrary code? In the rest of this section we'll go over examples of
unsafe behaviors, before formally capturing what stack-safe means
formally and discussing enforcement techniques from the literature
in later sections.

\subsection{Unsafe Variations}

@@

\rb{Intuition that the callee can do anything in principle, but the machine will
halt if it does not follow the calling convention, etc.?}

@@

In this section we describe our machine model. Our definition of stack
safety is generic and this section discusses an abstract interface to
our machine model and some kind of static or dynamic, software- or
hardware-enforced security monitor. We will refer back to the running
example of the previous section for concreteness, and then,
in~\cref{sec:micropolicies}, extend this model with a dynamic policy
enforcement mechanism in~\cref{sec:micropolicies} based on
\bcp{NickPaper,AndMaybeSomeOthers}.

@@

\leo{Other choices for observations? Where should that be discussed?}
\bcp{But in this paragraph we are describing the particular
choise that we make for our running example, no?  But we certainly want to
discuss this fully someplace!}
\leo{Here: Many observation choices. Next: Simplify:}
Mapping a conventional architecture to this abstract interface is
straightforward. For the running example of the previous section,
$\WORDS$ is the set of 64-bit integers, $\REGS$ is the set $\{r_0,
\cdots, r_{31}\} \cup \{\PCname\}$, and $\SP = r_2$. Observations are
the writes to the special $\rout$ location: if during an execution
step a word $w$ is written to $\rout$ then $\obs = w$, otherwise $\obs
= \tau$.

@@

\leo{Which of the machine, machine-observation, machine-policy,
  and machine-policy-observation traces should we distinguish
  in metavariables?}
\rb{Should we have notation for traces somewhere here?}

A trace over a type $\TRACEELEMS$ is a nonempty, finite or infinite sequence of
$\TRACEELEMS$s. We will mostly be concerned with sequences of pairs of machine states
and observations, called machine traces and denoted $\machT \in \MOTRACE$,
and with sequences of triples of a machine state, a policy state, and an
observation, denoted $\MPT \in \MPOTRACE$ and referred to as MP-traces. 
\rb{The way the definitions are set up, one may expect MPO, but\ldots}

@@

In our running example in Figure~\cref{fig:running-example-b} we can
see a depiction of such a trace: starting from an initial machine
state at the very top, each instruction constructs a new machine state
(the diff of which compared to the previous one appears in the
``Effect'' column) and the majority of observations are $\tau$ with
the exception of instructions 31 and 13 which are writes to $\rout$
and where the observations are $0$ and $60$ respectively.

@@

\begin{minipage}{.4\textwidth}
\judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
\end{minipage}
\begin{minipage}{.55\textwidth}
\judgmentthree{\(\mach_0 \stepstoobs{\obs} \mach_1\)}
              {\(\pol_0 \rightharpoonup \pol_1\)}
              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \machT\)}
              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT\)}
\end{minipage}
\rb{Shouldn't the conclusion of the first judgment use $\hookrightarrow_P$?
  And in the second judgment, wouldn't $MO$ be $MPO$, and the third premise $\hookrightarrow_P$?}

@@

\paragraph*{Observational Equivalence}

@@

>>>>

But what {\em is} this ``stack safety'' of which we speak? Precise formal
definitions are lacking in the literature; instead, the notion of safety
is usually defined negatively, by enumerating the bad things that a safe system
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
We are only aware of one, quite recent, attempt to frame a positive definition.
\citet{Skorstengaard+19} define stack safety as
the conjunction of two properties, {\em local state encapsulation} and {\em
  well-bracketed control flow}.
They formalize these properties by defining an abstract capability-based machine
that ``satisfies [stack safety] by construction.''
%
While this is an important step in the right direction, we believe
there is still room to craft a more abstract and less intensional
characterization of stack safety, so that we are able to answer
questions about other enforcement mechansims. That is the goal of this
paper.
\leo{I took a stab. I think it's still to dense, but captures the
  feeling of today's discussion I think}
\apt{still needs tuning}\bcp{Yes, I think we need to go into a
  little more detail on what their definition looks like in order to
  justify why an alternative definition might be valuable.  What
  questions does their presentation not answer? What uses does it not
  support? E.g., can it be enforced efficiently? (I guess so, by
  capabilities, though I think their capabilities are more powerful
  than what Cheri actually provides, while our policies are exactly
  what Nick needs.)  Why would it be good to look at stack safety
  through the lens of security?  (Because it is fundamentally a
  security mechanism?)}  \rb{BCP idea: contrast intrinsic
  vs. extrinsic property definition.  Remember to talk about spatial
  vs temporal safety}

Our main contribution in this paper is an alternative formal presentation of
local state encapsulation and well-bracketed control flow using the
conceptual and technical framework of language-based security.
%
The underlying notion behind local state encapsulation is the intention to
protect callers from callees: a callee should be not be able to access or
overwrite private data in the caller's stack frame. We formalize this
intuition as a security property with both confidentiality and integrity
components.
%
Intuitively, {\em stack confidentiality} promises that a callee's
observable behavior is not affected by memory outside its stack frame
(e.g. a caller's private data or uninitialized memory).  {\em Stack
  integrity} guarantees that, if a function call ever returns, the caller's
stack data at the point of return will be identical to what it was before
the call.  \bcp{We need to point out somewhere that we are only dealing with
stack safety, though we discuss (somewhere) how our property can be combined
with the heap safety property of [Azevedo et al]}\leo{Should this be
  here or in the assumptions as presented in the running example?}

@@

To provide evidence that our definitions capture the inituitive notion
of stack safety, we show that they rule out a range of exploitable bad
behaviors.\apt{``from the literature''??}  To assess whether they are
implementable, we compare them against the proposed capability-based
enforcement mechanism of \citet{Skorstengaard+19} \apt{we really
  should} and the existing tag-based enforcement mechanism of
\citet{DBLP:conf/sp/RoesslerD18}.\leo{Is this where we would rephrase this
  because we would only talk about it in related work?}

In summary, we offer the following contributions:
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
  Abstract formal properties capturing the
  intuitive concepts of local state encapsulation and well-bracketed
  control flow~(\cref{sec:lse-and-wbcf}) in a generic setting. These properties
  rule out common examples of stack memory abuse.
\item
  A case study of the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\cref{sec:enforcement}).
  This policy enforces our stack safety property, and also a (provably) stronger property that is easier to test.
  \apt{If we have no testing, do we even want to talk about the testing
    property?} \bcp{Not here; just in future work.  I guess we need to see
    whether we wind up with any results at all about testing.}
\item
  A refinement of the case study that considers the more efficient Lazy Tagging and Clearing policy
  of \citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:lazy}).  This policy does not enforce our property, but we give a
  variant of the policy that does enforce a (provably) weaker variant of our
  property.
%
  \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
    but the point we can make is still strong: To make the property easier
    to enforce, we can slightly weaken its guarantees; the weaker version is
    a bit more complicated to state but still retains the pragmatic
    protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
    depth-based tags, which do not enforce even our lazy policy. So we need to
    introduce a version of LT\&C that uses unique activation tags.}

@@

The example program appears in \Cref{fig:running-program}. It consists
of a {\tt main} function, written in some C-like high-level language,
that performs a call to some function {\tt f}, which in turns makes a
nested call to some function {\tt g} before returning, while
performing some simple arithmetic operations in the process.

@@

\newcounter{pcctr}
\setcounter{pcctr}{0}
%  \thepcctr & \stepcounter{pcctr} #1 & #2 & #3 & \text{#4}\\
\newcommand*{\row}[4]{
  \thepcctr & \stepcounter{pcctr} #2 & #3 & \text{#4}\\  
}
\newcommand*{\tracerow}[5]{
  #1 & #2 & #3 & #4 & \text{#5}\\
}
\newcommand*{\summary}[6]{
\[
\begin{array}{cccccc}
  \PCname = #1 & r_0 = #2 & \rra = #3 & \rsp = #4 & r_4 = #5 & r_5 = #6
\end{array}
\]
}
\newcommand{\negate}{\textrm{-}}

\colorlet{lgray}{gray!40}

\newcommand{\instrc}{lgray}
\newcommand{\outc}{white}
\newcommand{\fprivc}{cyan}
\newcommand{\fcallc}{green}
\newcommand{\gprivc}{yellow}
\newcommand{\gcallc}{orange}
\newcommand{\hprivc}{red}
\newcommand{\unusedc}{gray}

@@

  \begin{array}{c|c|c|c}
    PC & \mathit{Instruction} & \mathit{Effect} & \\

@@

        {Make the call (Jump and link)}

@@

%  \apt{How's that description for last line?}
%  \apt{Where did these RISC-V insruction formats come from? I think some of them are just wrong, e.g.
%    $\jal$ takes a PC-relative argument, not an absolute one, and the proper way to code a return to $\rra$ is $\jalr  ~ \rra ~ r_0$.
%  But also the instruction operand ordering is awkward. Why not use standard AT\&T syntax?}

@@

  \begin{array}{c|c|c|c}

@@

    \row{\jal ~ \rra ~ 40}{\jal ~ \rra ~ 40}{\PCname \leftarrow 40; ~ \rra \leftarrow \thepcctr}
        {Make the call to {\tt h} (jump-and-link)}

@@

\begin{figure}
\[
\begin{array}{c|c|c|c}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp - 2] \leftarrow r_4}{[107] \leftarrow 18}
      {Store $r_4$ as the return value}
  \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\rra \leftarrow 27}
      {Load return address}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 1}{\rsp \leftarrow \rsp - 1}{\rsp \leftarrow 108}
      {Restore the stack pointer of {\tt g}}
  \row{\jalr ~ \rra ~ \rra ~ 0}{\leo{??}}{\PCname \leftarrow 27}
      {Return to {\tt g}}
  \end{array}
\]

@@

\setcounter{pcctr}{27}
% H - Body
\vspace*{0.2em}
\[
\begin{array}{c|c|c|c}
  \row{\lw ~ r_4 ~ \rsp ~ 0}{r_4 \leftarrow [\rsp]}{r_4 \leftarrow 18}
      {Load return value to $r_4$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[105] \leftarrow 18}
      {Set {\tt w} to 18}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 106}
      {Deallocate {\tt h}'s call frame}
  \row{\lw ~ r_5 ~ \rsp ~ \negate 4}{r_5 \leftarrow [\rsp - 4]}{r_5 \leftarrow 0}
      {Load argument {\tt z} $r_5$}
  \row{\sw ~ \rout ~ r_5 ~ 0}{[\rout] \leftarrow r_5}{[99] \leftarrow 0}
      {Print {\tt z}}
  \row{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow r_4 + r_5}{r_4 \leftarrow 18}
      {Add $r_4$ and $r_5$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[103] \leftarrow 18}
      {Store $r_4$ as ther return of {\tt g}}
  \row{\lw ~ \rra ~ \rsp ~ \negate 2}{\rra \leftarrow [\rsp - 2]}{\rra \leftarrow 8}
      {Load return address}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 104}
      {Deallocate {\tt g}'s local state}
  \row{\jalr ~ \rra ~ \rra ~ 0}{??}{PC \leftarrow 8}{Return to {\tt main}}
\end{array}
\]
% Return to main
% Memory Layout
\begin{center}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{43.5em}{2em}{\SP}
\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{white}[{\makebox[0pt]{Output}}]%
\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
\memory{2}{\fcallc}[{\makebox[0pt]{{\tt g}'s call frame }}]%
\memory{6}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-51em}{0.75em}{0}
\MemoryLabel{-47em}{0.75em}{42}
\MemoryLabel{-43em}{0.75em}{0}
\MemoryLabel{-39em}{0.75em}{0}
\MemoryLabel{-35em}{0.75em}{18}
\\
\end{center}
% Register File
\summary{8}{0}{\thepcctr}{104}{18}{0}
%\vspace*{0.2em}
\setcounter{pcctr}{8}
\[
\begin{array}{c|c|c|c}
  \row{\lw ~ \rsp ~ r_4 ~ \negate 1}{r_4 \leftarrow [\rsp - 1]}{r_4 \leftarrow 18}
      {Load return value to $r_4$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[101] \leftarrow 18}
      {Set {\tt y} to the result}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 4}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 102}
      {Deallocate {\tt g}'s call frame}
  \row{\lw ~ \rsp ~ r_5 ~ \negate 2}{r_5 \leftarrow [\rsp - 2]}{r_5 \leftarrow 42}
      {Load {\tt x} to $r_5$}
  \row{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow r_4 + r_5}{r_4 \leftarrow 60}
      {Add {\tt x} and {\tt y}}
  \row{\sw ~ \rout ~ r_4 ~ 0}{[\rout] \leftarrow r_4}{[99] \leftarrow 60}
      {Print {\tt x + y}}
\end{array}
\]
\vspace*{0.2em}
% Return to main
% Memory Layout
\begin{center}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{31.5em}{2em}{\SP}
\memory{3}{\instrc}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{white}[{\makebox[0pt]{Output}}]%
\memory{2}{\fprivc}[{\makebox[0pt]{${\texttt{main}}_{\mathit{priv}}$}}]%
\memory{8}{\unusedc}%[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-51em}{0.75em}{60}
\MemoryLabel{-47em}{0.75em}{60}
\MemoryLabel{-43em}{0.75em}{42}
\\
\end{center}
% Starting Register File
\summary{\thepcctr}{0}{37}{102}{60}{42}
\caption{Detailed Execution Trace
  \rb{A suggestion for the description of $\jal$ and $\jalr$:
    call(function label) [if labels are added next to the entrypoints]
    or call(address of entrypoint) and return()?}
}
\label{fig:running-trace-b}
\end{figure}

To show a trace of the execution of this program, we need a low-level
view of both the code and the machine it runs on. While the concrete
implementation of our enforcement mechanism (\Cref{sec:enforcement})
operates on a full conventional RISC-V architecture (using the FORVIS
simulator\leo{cite}), we make a few simplifying assumptions here to
ease the exposition.

First of all, we ignore the heap. We are mainly concerned with stack
interactions so that will be the main focus; integrating a heap in the
style of \citet{AzevedoDeAmorim17} is a natural extension we discuss
in~\Cref{sec:future}.
%
We also abstract away low-level details of memory layout and present a
machine that is word-addressed, with instructions starting from
location 0, a dedicated output at location 99, and lay the stack from
location 100 upwards. \leo{Explain more? How to connect to realistic
  one? Observations? DMA? Interrupt handlers?}
%
A sketch of the memory layout at that point can be found in the top of
\cref{fig:running-trace-b}: instructions for {\tt main} begin at 0,
the dedicated output is uninitialized, while the (also uninitialized)
stack begins at 100. Directly below the layout, we summarize the parts
of the register file that are relevant to the example.
%
The rest of \cref{fig:running-trace-b} details the execution of the
program, showing the machine state at call and return points, as well
as describing the full effect of each instruction in detail.
\leo{Expand upon this?}

@@

Stack safety is meant to protect a caller from a malicious callee's
attempts to access private data or highjack its control flow. Our
attacker model reflects this intention: we assume that the attacker
can run arbitrary code as a caller and the enforcement mechanism
should protect against that. Hardware and timing attacks are out of
scope. \leo{Anything else to add here?}
\leo{Maybe a good place to discuss calling conventions and that they
  are essentially enforced by the policy or sth?}
\rb{Intuition that the callee can do anything in principle, but the machine will
halt if it does not follow the calling convention, etc.?}

\subsection{Unsafe Variations}
\sna{Variations here collides with state variations. I propose substitutions.}
\leo{I agree there is a collision, not sure about ``substitutions''}

The execution trace shown in \cref{fig:running-trace-b} is an example
of a ``stack-safe'' execution: each process interacts only with the
registers, its local stack space and its call frame. In the rest of
this section we'll go over examples of unsafe behaviors, before
formally capturing what stack-safe means formally and discussing
enforcement techniques from the literature in later sections.

@@

@@

In this section we describe our machine model. Our definition of stack
safety is generic and this section discusses an abstract interface to
our machine model and some kind of static or dynamic, software- or
hardware-enforced security monitor. We will refer back to the running
example of the previous section for concreteness, and then,
in~\cref{sec:enforcement}, extend this model with a dynamic policy
enforcement mechanism based on
\citet{DBLP:conf/sp/RoesslerD18}.

@@

The choice of observations is an interesting one. A first approach
would be to consider any change to the registers or memory observable.
However, as some parts of the stack will be considered ``private''
this choice won't work. A relaxation of this to consider all changes
to ``non-private'' data observable could work in principle, but
doesn't allow for an interesting weaker-but-more-efficient enforcement
mechanism. We'll come back to this point in \Cref{sec:lazy}. On the
other hand, we could use a particular location as an abstract output
of our system (e.g. through DMA). This is the approach we chose for
our running example to keep things simple. For that, $\WORDS$ is the
set of 64-bit integers, $\REGS$ is the set $\{r_0, \cdots, r_{31}\}
\cup \{\PCname\}$, and $\SP = r_2$. Observations are the writes to the
special $\rout$ location: if during an execution step a word $w$ is
written to $\rout$ then $\obs = w$, otherwise $\obs = \tau$.

@@

A trace over a type $\TRACEELEMS$ is a nonempty, finite or infinite sequence of
$\TRACEELEMS$s. We will mostly be concerned with sequences of pairs of machine states
and observations, called machine traces and denoted $\machT \in \MOTRACE$,
and with sequences of triples of a machine state, a policy state, and an
observation, denoted $\MPT \in \MPOTRACE$ and referred to as MP-traces.

@@

In our running example in \Cref{fig:running-trace-b} we can
see a depiction of such a trace: starting from an initial machine
state at the very top, each instruction constructs a new machine state
(the diff of which compared to the previous one appears in the
``Effect'' column) and the majority of observations are $\tau$ with
the exception of instructions 31 and 13 which are writes to $\rout$
and where the observations are $0$ and $60$ respectively.

@@

\begin{minipage}{.4\textwidth}
\judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_p \mpostate{\mach_0}{\pol_0}{\tau}\)}
\end{minipage}
\begin{minipage}{.55\textwidth}
\judgmentthree{\(\mach_0 \stepstoobs{\obs} \mach_1\)}
              {\(\pol_0 \rightharpoonup \pol_1\)}
              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \MPT\)}
              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \MPT\)}
\end{minipage}

@@

\paragraph*{Observational Similarity}

@@

\medskip
